---
title: "Tutorial4 Visualizations (without code)"
author: "Caroline Kjær Børsting"
date: "9/23/2021"
output: html_document
---


In this tutorial, you learn the basics of how to do visualizations in R, such as how to create ggplot objects, change colors, facets, legends etc. In addition to the basic stuff, you will also get to play around with some of my favorite plotting functions in R; for example how to plot heatmaps on top of real geographical maps, create time series plots which move as a function of time and how to turn boring scatter plots into GIFs. These visualization methods will hopefully provide us with some great insights about app reviews, stock prices, police killings and worldwide GDP. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("pacman") #install the pacman package first if you haven't installed it before. 
install.packages("pacman")

#We will use different functions from a lot of different packages. The easiest way to install and load multiple packages at once is to use the p_load function from the pacman package:
pacman::p_load(
  ggplot2, #for plotting graphics
  ggmap, #for plotting maps
  RColorBrewer, #provide pretty colour palettes
  lubridate, #for converting time variables into date format
  hrbrthemes, #for more style themes
  data.table, #for working with data.table
  dplyr, #for count function
  tidytext, #for text mining functions such as unnest()
  wordcloud,
  tm, #for removing stopwords
  ggdist, #for raincloud plots
  tidyquant, #for theme_tq
  prismatic, #for color handling in plots
  gganimate, #for animated plots
  gifski #for rendering animated plots as GIFs instead of single frames
)
```

##Apple stock prices
The first dataset we will work with today is Apple stock prices from 1980-2021. Data is kept daily and presented in USD currency.

###Loading data:
```{r}
#Let's start by loading the data using the read.csv() function: (hint: the dataset is caled "AppleStockPrices.csv")
AppleStockPrices <- read.csv("AppleStockPrices.csv")
  
#Then we can see what the data looks like by displaying the headers: (hint: use the head() function)
head("AppleStockPrices")
  
#As you could see, this dataset contain timeseries of stock prices over a period of 21 years. We will therefore treat the data as a timeseries. When working with timeseries, it is important that our time variable is in the correct format. We can check what the format of our variables is by running the following command which displays the structure of our data: (hint: use the str() function)
str(AppleStockPrices)
  
#The second column displays the format of each variable. As you can see, all variables are numeric ("num") except the first variable "Date" which is a character string ("chr"). The "Date" variable will be our time variable, and we therefore need it to take the format "Date" instead of "chr". 

#Luckily, the "ymd" function from the "lubridate" package" allows us to do this very easily: (Hint: ymd is short for year-month-date format. Therefore, insert the Date variable from our dataset into the ymd() function)
library(lubridate)

AppleStockPrices$Date <- ymd(AppleStockPrices$Date)

#So how does it look now? (hint: use the str() function)
str(AppleStockPrices)
  
#Exactly as we want it to, so now we are ready to start plotting!
```

###Timeseries plot
We will base most of today's plots on the "ggplot2" package which you can read more about by typing "ggplot2" in the Help tab. 
```{r}
#Let's start out by by making a simple lineplot with time on the x-axis and opening stock price on the y-axis: (hint: insert the dataset and an aesthetics argument into the ggplot() function. Use Date as our x-variable and Open as our y-variable. Add geom_line() on a new line)
library(ggplot2)
AppleStockPrices %>% 
  ggplot(aes(Date, Open)) + geom_line()

#We can change the axis names by adding labels: (hint: add a labs() argument on a new line and define new labels as x = "some name", y = "some name".)
AppleStockPrices %>% 
  ggplot(aes(Date, Open)) + 
  geom_line() + 
  labs(x = "Date", y = "Opening Stock Price")

#And add a title: (Hint: insert a "tittle" argument inside the above labs argument)
AppleStockPrices %>% 
  ggplot(aes(Date, Open)) + 
  geom_line() + 
  labs(x = "Date", y = "Opening Stock Price", title = "Stock Prices Apple")

?labs

#We can also rotate the text along the x-axis, if we think that this looks better: (Hint: Add a new line with a theme() function. Insert an element_text argument, set angle to 60 degrees and hjust to 1 for a horizontal rotation of 60 degrees. Set this to equal axis.text.x.)
AppleStockPrices %>% 
  ggplot(aes(Date, Open)) + 
  geom_line() + 
  labs(x = "Date", y = "Opening Stock Price", title = "Stock Prices Apple")

?theme

#And we can change the color of the line, if we think that this looks better: (hint: insert a "color" argument inside the geom_line function and find a nice-looking colorcode here: http://sape.inf.usi.ch/quick-reference/ggplot2/colour)


#Or add a straight line at some level, if we think that this adds something to the story: (hint: add a new line with the geom_hline() function. Set yintercept to some random value. You can also specify a color and size of this new line)
>
```

###Dual timeseries charts
```{r}
#We can also add one more line representing one more variable to our lineplot. Fx we could plot the stock volume together with the opening price: (hint: again, insert your dataset to a ggplot() function, and define only the x-value inside the aesthetics argument. Then add a new line with a geom_line() where you insert an aesthetics where you only define a y-value and give it a color. Then you add another line with a geom_line() where you insert an aesthetics where you define another y-value and give it another color. You can also add labs() and theme() arguments as above)
>
  
#This looks a bit weird, given that the two variables are represented on quite different scales. 

#We can help this a bit by scaling one of the variables to make it look a bit prettier: (hint: use the scale() function, insert the Volume variable and set the center argument to false.)
>

#How does it look now? (hint: plot the same plot again)
>
  
#Better but not awesome. Probably, it is just not very meaningful to plot stock price and stock volume together, so let's just leave it there. But I hope this worked to illustrate how easy it is to plot several variables on top of each other. 
```

###Animated timeseries 
```{r}
#The "gganimate" package allows us to turn our timeseries plots alive! 

#First, we must create our ggplot object and add a line- and point geom. (hint: use ggplot() again, set aesthetics to be x=Date, y=Open again, add a geom_point(), add a geom_line, add a new line with a transition_reveal() function from the gganimate package and insert the Date variable iside this function. Save it as some name so that we can later call it.)
>
  
#we use the "gifski" package to render our object into a moving GIF and not just seperate pictures. (Hint: use the animate() function, insert your gg-object from above, and use the gifski_renderer() function to render it as a GIF)
>
  
#Now, let's add a title to the gif. I would like to have both a normal title and a subtitle that contains the given year. Currently, we don't have a variable that contains only the year; all we have is a variable containing the date. We can, however, use the format() function to isolate the year from the date: (hint: insert your Date-variable into the format() function at set format to equal "%Y" because we only want the year.)
>
  
#And then insert it as another line in our gg-object: (hint: add a labs() function, and set the subtitle argument to equal "Year: {format(frame_aling, "%Y")}. The combination of the format() function and frame_along allows you to only plot a given year at a given timepoint)
>

#Again, we must render our gg-object: (hint: as before, use the animate() function, insert the gg-object and set renderer to equeal the gifski_renderer() function)
>
  
#Besides making the timeseries move as the stock price changes, you can also make the axes move and keep the timeseries stable as the stock price changes. We do this by using the view_follow() function from the gganimate package: (hint: all you have to do is adding a view_follow() function on a new line)
>
  
#Again we must render our gg-object: (hint: as before, use the animate() function, insert the gg-object and set renderer to equeal the gifski_renderer() function)
>
```

##App review sentiment:
The second dataset that we will work with today contains reviews scraped from the Google Play Store for the ecommerce company Nykaa, ending August 2021. The dataset cotains the actual review text as well as a sentiment score binning, where only reviews with 5 star rating are considered POSITIVE. Those with 3 and 4 star rating are considered NEUTRAL and everything else is NEGATIVE.

###Loading data
```{r}
#First we will load the dataset "nykaa_review_sentiment.csv" using the read.csv() function. 
>
```

###Basic plots
```{r}
#Having loaded the dataset, we can now take a look at the data to see what we're dealing with. 

#First of all, we can extract some insights from the sentiment_label variable. For example, it would be interesting to know how the reviews are distributed in terms of sentiment. To do this, we must first figure out how many reviews of each class the dataset contains. The data.table package offers a neat way of extracting this information by creating an aggregated dataset. 

#In order to use the functions from the data.table package, we must first turn out "sentiment" dataframe into a data.table: (hint: use the as.data.table() function and save it in the same name)
>

#Now we can create an aggregated dataset that contains the number of reviews of each of the "sentiment_labels" classes: (hint: use hard-brackets [ ] to specify that you want to make a subset of the data.table. Insert a list() function wherein you create a new variable that sums the number of elements of earch class in the "sentiment_labels" varialbe. Save it as sentiment_agg).
>
  
#We can then plot the sentiment labels from the the aggregated sentiment dataset: (hint: use ggplot(), insert the aggregated dataset, set x=sentiment_labels, y=sum inside the aesthetics argument, add a geom_histogram() where you set stat = "identity)
>

#And change size of the bins: (hint: add a "width" argument inside the geom_histogram)
>

#And the colors: (Hint: add a "fill" argument inside the geom_histogram)
>

#And we can add a title and labels to the x and y axis: (hint use the labs() function)
>

#And change the theme, i.e. appearance of the plot: (hint: Use fx the theme_linedraw() function)
>

#And change the size of the title: (hint: use the theme() function as set plot.title to equal element_text(size = some number))
>
```

###Wordclouds
A wordclould is a graphical representation of word frequency of a given text which give greater prominence to words that appear more frequently than others. Thus, wordclouds offer nice and intuitive insights about what is being "talked about" in any given text. 
You will learn much more about the linguistic considerations when preparing your text data for text mining analyses such as tokenization, lemmatization, etc at the lecture with Michal Zaggl. Therefore, we will not delve too much into these details today. 
```{r}
#First we must create a new dataset that contains words seperately: (hint: use the unnest_tokens() function, group_by() function, mutate() function and ungroup() function). 
>

#Many words appear in multiple reviews, so let's sort them so that they all only appear once: (hint: use the count() function and set sort to be true)
>

#Now we are ready to plot our wordcloud: (hint: use the wordcloud() function, specify the words, freq, min.freq, max.words, random.order, rot.per, and colors arguments)
>

#As you can see, many of the words that appear most frequently does not carry much insight, such as "the", "and", "is", "for", "a". In linguistic terms, these words are called "stop-words" and are usually always removed in text mining analyses, since they don't carry much meaning. 
#The "tidytext" package contains a lexicon called "stop_words" which contains the most common stop-words in the English language. (Hint: We can remove these words from our "sentiment_words" dataframe using the anti_join function from the "dplyr" package) 
>
  
#Again, let's sort them so that they all only appear once: (hint: again, use the count() function and set sort to be true)
>
  
#And plot our wordcloud: (hint: again, use the wordcloud() function, specify the words, freq, min.freq, max.words, random.order, rot.per, and colors arguments)
>

```

##Police killings
The next dataset contains information of police killings in the United States in 2015, and originates from the Guardian via Census. 

###Loading data
```{r}
#First we load our data "police_killings.csv" using read.csv():
>
```

###Basic plots
```{r}
#First let's look into how the victims were distributed in age: (hint: use ggplot, only specify an x-variable of age in the aesthetics and add a geom_density() function)
>

#We get an error message saying: "Error: Discrete value supplied to continuous scale"

#R thus complains about our variable "age" not fitting the geom_density() function whih is a density curve on a continuous scale. 
#Let's therefore take a look at how our variables are structured: (hint: use the str() function)
>
  
#As you can see, our "age" variable is currently in the "chr" format, i.e. a character variable. We can transform it to a numeric variable using the as.numeric() function: (hint: if passing our variable directly to the as.numeric() function, you will see that it doesn't recognize the numbers that the variable already holds in the chr format, and instead creates new numerical values for every character value. We therefore have to tell it first to treat it as a character and then as a numeric.)
>

#And now we can try to plot it again: (hint: again, use ggplot, only specify an x-variable of age in the aesthetics and add a geom_density() function)
>

#Now that is it working, let's see if we can make it look a bit prettier! First we can change the color of the filling/what is under the density curve: (hint: add a "fill" argument inside the geom_density() function)
>

#Then we can change the color of the curve line: (hint: add a "color" argument inside the geom_density() function)
>

#We can also make the density plot slightly transparent by adding an alpha argument: (hint: add an "alpha" argument inside the geom_density() function)
>

#And we can change the theme for a more streamlined look: (hint: add fx a theme_bw() function on a new line)
>

#And we can add a title and x and y labels: (hint: add a labs() function on a new line)
>

#And finally, we can change the x scale to be within a certain range: (hint: use the xlim() function)
>

#Another cool feature in ggplot2 is facet-wraps! Facets allow you to split your plots across different variables. For example, we could check whether there is a difference in age distribution of the deceased across ethnicities: (hint: add a facet_wrap() function on a new line, and insert "~raceethnicity inside the function)
>

#Instead of looking at the age of the deceased, we could also look into whether there is an even distribution of ethnicity in American police killings: (hint: substitute the x variable in the above plot with "raceethnicity and remove the facet_wrap() statement)
>

#Clearly there is an over-representation of White Americans. However, given that 60 % of Americans are Caucasians and only 13 % are African American, the White Americans might not be so over-represented in the statistics after all. Food for thought...

#You can further split your results on more variables. For example, we could fill the bins from the above plot with information of whether the deceased was armed or not: (hint: add a fill = armed argument to the aesthetics in the above plot to fill out the bars based on whether the deceased were armed or not)
>
#However, this often comes with a compromise of readability, as you can see. 

#Finally, we can also plot the distribution of ethnicity as a pie chart/circle diagram. In order to do this, we must first extract some summary information which we can save directly in a dataframe: (hint: use the data.frame() function and use the summary function inside. You may have to wrap your variable in a as.factor function)
>
#As you can see, the races are saved as rownames. We don't want that. Instead, we want them to be saved as a seperate column. (hint: The "tibble" package has a handy function for that called rownames_to_column(). Insert the summary dataframe and specify a name for the new column)
>
  
#And now we're ready to create our pie chart using a combination of the geom_bar() and coord_polar() functions. The scale_fill_brewer() function comes from the "RColorBrewer" package and allows you to choose the color palette you like the most. Google "RColorBrewer palettes" to find more color codes. (hint: use ggplot(), in the aesthetics argument, set x to be empty, y to be the summary variable from above, and fill to be the name of the new race column. Add a geom_bar() function where stat = "identity" and width =1. Add a coord_polar() function where you set "y" and start=0. Finally, add a scale_fill_brewer() function and set palette to equal "some color code".)
>
  
```

###Heatmap
R also offers a fairly simple way of creating geographical heatmaps based on longitude and latitude coordinates using the "ggmap" package. In the below code chunk, we will create a heatmap laid over Northern America to get a graphical representation of where most police killings take place.  
```{r}
#First we must define the bounds of the to be mapped country. United States in our case. Country-wise bound codes can be found on google. (Hint: the bounds of USA is left = -125, bottom = 25, right = -65, top = 50)
>
  
#We configure the map, bounds, zoom and style, via the get_stamenmap() function in the "ggmap" package. (Hint: We feed the function with the map-bounds we defined above so that the function knows to search for the stamen tiles that fits the United States, with a zoom-in level and with a maptype argument which defines whether the background should be terrain, toner etc.)
>
  
#Then we render the map using the ggmap() function from the "ggmap" package based on the above coordinate map which is a ggmap-object. We can now plot a visual representation of our police killing data based on two-dimensional kernel density estimation using the stat_density2d() function. The "..level.." argument tells stat_density2d() function to reference the given level of police killings in the dataset, so that more filled-out locations and less transparent locations on the plot reflects more police killings. Finally, we ask the ggmap() function to plot it as a polygon:
>

#We can also play around with the color palette (Hint: use the scale_fill_gradientn() function for creating n-color gradients. The brewer.pal() is most commonly used to create thematic maps like this. By wrapping it in a rev() statement, you reverse the order of the colors.)
>

#And change the stylistic theme of the plot, such as giving it a black-and-white theme: (HInt: use the theme_bw() function)
>
  
```

##GDP data
The last dataset that we will be working with today contains information about GDP per capita and life expectancy across different countries.

###Loading data
```{r}
#As per usual, we start by loading the data: (hint: use the read.csv() function. Dataset is called "gdp_data.csv")
>
```

#Basic plots
```{r}
#First, we can make a simple line plot that smoothes over conditional means. (Hint: use the geom_smooth() function. Set the aesthetics to be x = gdpPercap, y = lifeExp. By setting the smoothing method to "lm", we state that we expect the simplest type of relationship: a linear relationship.) 
>
  
#If you change the method to "auto", R chooses a smoothing method automatically (based the size of the largest group across datapoints): (Hint: substitute the "lm" statement with "auto")
>
  
#Now we can see that the relationship might actually be closer to e.g. a gamma distribution, where life expectancy strongly increases as GDP increases, but eventually flattens out and actually starts decreasing for the very rich countries. 

#We can also do a boxplot of life expectancy across continents. (Hint: use the geom_boxplot() function. Change the x aesthetic to "continent".)
>
            
#Or we would make a scatterplot of GDP vs. life expectancy for all data points. (Hint: use the geom_point() function. Change the x aesthetics back to gdpPercap.)    
>
  
#Let's try to make the above plot more readable and better looking by adding a few arguments! First, we can color code the data points based on continent by adding a color argument in our aesthetics: (Hint: set color to equal "continent" inside the aesthetics function)    
>

#We can also make the data points differ in size based on the size of the population for the given country: (Hint: set size to equal "pop" inside the aesthetics function)    
>
  
#Now we have a rather bastiant looking legend that explains the sizes of the dots, which we don't really need. Let's try to remove this. (Hint: add a guides() argument and set size to "none")
>

#We can change the position of the legends explaining data point color using the theme() function: (Hint: add a legend.position argument inside the theme() function and set it to fx "bottom".)
>

#And we can make the datapoints slightly transparent to better tell them apart (Hint: add an "alpha" argument inside the geom_point() function and set it to a value between 0-1)
>

#And we can again change the stylistic theme of the plot: (Hint: add fx the theme_minimal() function)
>

#And finally we can also change the title labels on the axes of our plot: (Hint: use the labs() function and add title, x and y arguments.)
>

#To me, this plot now looks much prettier than the original black-and-white version. Yay for pretty plots! However, there is still one thing that we can add to really make the plot come alive: actually making it live! (Hint: we can use the transition_time() function to only show data points for a given year right after each other, so that it looks like the plot is moving across time (in the period 1952-2007). Add it as a new line on the plot.)
>

#And render the GIF using gifski. (Hint: Use the animate() function, pass the name of the animated plat and set renderer to use the gifski_renderer() function to render the GIF)
>
  
#We can also add a subtitle that tells us the current year displayed: (Hint: use the labs() function, and set subtitle to equal "Year: {format(frame_time)}" to extract the name of the current year)
>

#And again render. (Hint: Use the animate() function, pass the name of the animated plat and set renderer to use the gifski_renderer() function to render the GIF)
>

#anim_save("gdp.gif", pp) #if you want to save this pretty lil' GIF, you can use this command!
```

###Raincloud plots
Hopefully, you have gotten a sense of how plots can be build up by first creating a gg-object using the ggplot() function and afterwards adding geoms, themes etc. A graph type that is currently gaining popularity is the Raincloud plots because it carries a lot of information at once. Raincloud plots usually consists of a density plot or half of a violin plot that is flipped 90 degrees so that it is "lying down" and thus looks like a cloud. Below the cloud is a scatter plot consisting of individual data points scattered along the x-axis so that the data points look like rain drops. Often, a boxplot is laid on top of the scatter plot for a visualization of summary statistics (mean, upper- and lower quantiles, outliers). 
This way, the raincloud plot provides the user with information both about individual observations and patterns among them (such as clustering), and overall tendencies in the distribution. This is important, because even a boxplot plus raw data may hide bimodality or other crucial facets of the data.
```{r}
#Hint: we use the ggplot() function in which we set aesthetics to be x = continent, y = lifeExp. We add a density plot using the stat_halfeye() function, a boxplot using geom_boxplot(), scattered data points using geom_jitter. On top of these basic elements, we use the coord_flip() function to make the plots horizontal, and we can additionally use the coord_cartesian() function to specify how the plot should be displayed, the labs() function to change labels, and the theme() function to change the visual appearance. See individual specifications in the Rmd file with code - this becomes to lengthy :-) 


